---
title: "metagear"
author: "Annemarie van der Marel"
date: "18/04/2021"
output: html_document
---

metagear package allows to
- Initialize a dataframe containing bibliographic data (tile, abstract, journal) from multiple study references.
- Distribute these references randomly to two team members.
- Merge and summarize the screening efforts of this team.

http://lajeunesse.myweb.usf.edu/metagear/metagear_basic_vignette.html
https://cran.r-project.org/web/packages/metagear/metagear.pdf

```{r setup, include=FALSE}
# first load Bioconductor resources needed to install the EBImage package 
# and accept/download all of its dependencies
#install.packages("BiocManager"); 
#BiocManager::install("EBImage")
#install.packages("metagear")

# then load metagear
library(metagear)
library(dplyr)
```

## distribute references


```{r load bib dataset}

# load a bibliographic dataset with the authors, titles, and abstracts of multiple study references 
refs <- read.csv("savedrecsWOS_2021-04-18.csv")
# display the bibliographic variables in this dataset
names(refs)

# change naming now or later
refs_selectVar <- refs %>% 
  dplyr::select(Authors, Publication.Year, Article.Title, Volume, Start.Page, End.Page, DOI, Abstract) 
example.names.metagear <-    c("AUTHORS"  "YEAR"     "TITLE"    "JOURNAL"  "VOLUME"   "LPAGES"   "UPAGES"   "DOI"      "ABSTRACT")

# prime the study-reference dataset
theRefs <- effort_initialize(refs) # adds 3 columns: unique key per ref, empty reviewer column, and a screening effort column 'include'
# display the new columns added by effort_initialize 
names(theRefs)



# randomly distribute screening effort to a team
theTeam <- c("Alex", "Alexis", "Annemarie","Catherine", "Connor", "Lauren", "Lisa", "Manuch", "Marya")
# randomly distribute screening effort to a team,
# but also saving these screening tasks to separate files for each team member
theRefs_unscreened <- effort_distribute(theRefs, reviewers = theTeam,  save_split = TRUE) # add effort (e.g. = c(20, 80))) to delegate screening effort unevenly
theRefs_unscreened[c("STUDY_ID", "REVIEWERS")] # display screening tasks

list.files(pattern = "effort")


```

## Screening abstracts of references 

Screening interface
no: to exclude study
maybe: more information is needed to determine inclusion
yes; include study

```{r screening abstracts}
# initialize screener GUI
abstract_screener("effort_Annemarie.csv", 
                  aReviewer = "Annemarie",
                  titleColumnName = "Article.Title",
                  abstractColumnName = "Abstract")


```
## Merge screening efforts

```{r merge screening effort}
# WARNING: will merge all files named "effort_*" in directory
theRefs_screened <- effort_merge()
theRefs_screened[c("STUDY_ID", "REVIEWERS", "INCLUDE")]
theSummary <- effort_summary(theRefs_screened)
```



## Generating PRISMA plots

```{r Generating PRISMA plot}

phases <- c("START_PHASE: # of studies identified through database searching",
            "START_PHASE: # of additional studies identified through other sources",
            "# of studies after duplicates removed",
            "# of studies with title and abstract screened",
            "EXCLUDE_PHASE: # of studies excluded",
            "# of full-text articles assessed for eligibility",
            "EXCLUDE_PHASE: # of full-text articles excluded, not fitting eligibility criteria",
            "# of studies included in qualitative synthesis",
            "EXCLUDE_PHASE: # studies excluded, incomplete data reported",
            "final # of studies included in quantitative synthesis (meta-analysis)")

thePlot <- plot_PRISMA(phases)
# PRISMA plot with custom layout
thePlot <- plot_PRISMA(phases, design = c(E = "lightcoral", flatArrow = TRUE))
```






