---
title: "import_networks"
author: "Annemarie van der Marel"
date: "11/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(igraph)
library(ggplot2)
library(einet)
library(EloRating)
library(tidyverse)
```

```{r}
#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}
```



# Import wood ants networks
Burns et al. 2020

In this folder there are six csv files that have all of the data for the 
exclusion experiment and previous data not used in the analysis:

1. conditions.csv     - This file has the conditions that each colony belongs to in the experiment
2. edges.csv          - This file has the edges (trail) data for all maps taken for the
		    	experiment and previous work on this system including the two nodes
		    	that the edge connects (undirected), the colony, the date and the
		    	the strength of each connection
3. focal_trees.csv-     This file has the focal trees for each colony in the experiment
4. nest_locations.csv - This file has the locations (x/y) for all the nests in the 
			experiment
5. nests.csv	      - This file has the nest data for all maps taken in the experiment 
			and previous work on this system including the node ID, colony,
			date and size of nest
6. trees.csv	      - This file has the tree data for all colonies in the experiment and
			previous work on the system including the tree ID, colony, species 
			and locations of the trees

# import data
```{r wood ants}
#path_in <- "./data/data Burns_ants food removal/"
#path_out <- "./analysis"  # where do we want to save the networks?

nests<-read_csv("../data/data Burns_ants food removal/nests.csv")
id_list<-unique(nests$ID)


conditions <- read_csv("../data/data Burns_ants food removal/conditions.csv") # ../ will go one folder back
edges <- readr::read_csv("../data/data Burns_ants food removal/edges.csv")

edges<-left_join(edges, conditions)

View(edges)
head(edges)
glimpse(edges)
unique(levels(as.factor(edges$colony)))
unique(levels(as.factor(edges$condition)))
unique(levels(as.factor(edges$date)))
unique(edges$from)
unique(edges$to)
```
Time points: "For this study we mapped each colony at four time points: (a) before we installed the collars, in August 2016; (b) after we installed the collars, in June 2017; (c) late in the season, in August 2017; and (d) after the first overwinter since the collars were installed, in June 2018. Here we use the August 2016 and August 2017 time points to assess changes to colonies and nests."

Are we using the same time points?


# number of networks (undirected)
- condition (2)
- period (2)
- colony (9)

# get data ready

```{r}
sub_edges<-edges %>% filter(date %in% c("22/08/2016","22/08/2017"))
sub_edges$period[sub_edges$date=="22/08/2016"]<-"before"
sub_edges$period[sub_edges$date=="22/08/2017"]<-"after"

colonies <- unique(levels(as.factor(sub_edges$colony)))

sub_edges$id <- paste0(sub_edges$colony, "-", sub_edges$period)

head(sub_edges)
glimpse(sub_edges)

```


First, try on one network
```{r}
# subset on 1 colony
sub_edges1<-filter(sub_edges, colony=="2b", period=="after")

graph_woodants<-graph_from_data_frame(sub_edges1)
graph_woodants
print(graph_woodants)
plot(graph_woodants)

```
# loop to get network and network metrics
```{r}
colony_timepoint<-unique(sub_edges$id)

 all_nets = list()  # empty list for networks
 
 #make empty dataframe to fill
 
 node_summary <- data.frame(focal.m=character(), 
                      degree=numeric(), 
                      eigenvector_centrality=numeric(),
                      betweenness_centrality=numeric())
 
 
network_summary <- data.frame(focal.m=character(), 
                      n_nodes=numeric(), 
                      density=numeric(), 
                      normal_ei=numeric())
                     #apl=numeric(),
                     #ei=numeric())
                     #)
                                    
  i=1

for (i in 1:length(colony_timepoint)) {
  # as a check and to know where in the loop we are 
  focal.m <- colony_timepoint[i]
  print(focal.m) 
  
  # subset the dataframe per combi
  focal.data <- subset(sub_edges, id==focal.m)
  head(focal.data)
  
  ## convert to network format
  graph<-graph_from_data_frame(focal.data, directed = F)
graph 
  
 #plot 
  plot(graph, 
       edge.width=E(graph)$weight,
       #vertex.color="grey",
       vertex.label="",
       edge.color=alpha("black", 1),
       vertex.size=20,
       layout=layout_in_circle,
       main=focal.m)


#### individual-level : Betweenness Centrality, Eigenvector Centrality, and Degree ###

  # degree
  degree <- igraph::degree(graph)
  
  # betweenness centrality -> standardized
  between <- igraph::betweenness(graph, directed = F,
                                      weights = E(graph)$weight)
  betweenness_centrality <- (between-min(between))/(max(between)-min(between))
  
  
  
# eigenvector centrality
eigenvector_centrality<-igraph::eigen_centrality(graph, directed = F)$vector


# pool individual
pool_id <- cbind.data.frame(focal.m,
                            degree,
                            betweenness_centrality, 
                            eigenvector_centrality)

node_summary <- rbind(node_summary, pool_id)
                        


#### network level (N Nodes, Density, Effective Information) ####

# number of nodes
  n_nodes <- length(V(graph))

# density =  number of observed dyads/total possible dyads
#density<-ecount(graph)/(vcount(graph)*(vcount(graph)-1)) #for a directed network
density<-igraph::edge_density(graph)

# average path length
apl <- igraph::mean_distance(graph) #average.path.length(graph.ref.crowd) 

# efficiency
library(einet)
ei <- effective_information(graph, effectiveness = FALSE)
normal_ei <- ei/log2(n_nodes) # normalized value to control for network size

# betweenness_centrality <- mean(igraph::betweenness(graph, directed = F, weights = E(graph)$weight))
  # eigenvector_centrality<-mean(igraph::eigen_centrality(graph, directed = F)$vector)

#Find proportion unknown relationships, a measure of sparseness
#prunk <- EloRating::prunk(matrix)
#prunk.pu <- as.numeric(prunk[1])
#prunk.dyads <- as.numeric(prunk[2])

## pool network data
pool_network<- cbind.data.frame(focal.m,
                      n_nodes,
                      density, 
                      normal_ei)

network_summary <- rbind(network_summary, pool_network)
     
# pool networks
  graph <- as.list(graph)
  names(graph) <- focal.m
  all_nets[[focal.m]] <- graph

                   
}
 
all_woodants_networks<-all_nets


```


```{r save output}
# compile node output and save 
head(node_summary)

node_output <- node_summary %>% 
  tibble::rownames_to_column() %>% rename(ID=rowname) %>% 
  separate(focal.m, c("replicate", "treatment")) %>%
  mutate(study="burnsetal2020", species="wood ants") %>%
  select(study, species, treatment, replicate, everything())
head(node_output)

write.csv(node_output, file = "../analysis_results/node_output_woodants.csv")



# compile network output and save 
head(network_summary)

network_output <- network_summary %>% 
  #tibble::rownames_to_column() %>% rename(nodeID=rowname)
  separate(focal.m, c("replicate", "treatment")) %>%
  mutate(study="burnsetal2020", species="wood ants") %>%
  select(study, species, treatment, replicate, everything())
head(network_output)

write.csv(network_output, file = "../analysis_results/network_output_woodants.csv")


# save network edgelist or matrix as csv
head(sub_edges)
glimpse(sub_edges)

write.csv(sub_edges, file = "../analysis_results/edgelist_woodants.csv")



 
```





