---
title: "import_networks"
author: "Annemarie van der Marel"
date: "11/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(igraph)
library(ggplot2)
library(einet)
library(EloRating)
library(geosphere)
```

```{r}
#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}
```



# Import tree lizards networks
Lattanzio & Miles 2014

It's not a before and after a perturbation. Instead, spatial networks are compared between different sites differing in the frequency of a disturbance. 
We could use the non-burned site as control (before treatment) and high burn site as after treatment.  

 'morph_networks'
Contains data for morphology, body condition (svl), parasite load, bite mark, perch use frequency, network construction (lat, long data) and analysis (degree, eigenvector centrality data), and morph frequency analyses. 

    'lat' and 'long' used to estimate spatial network at each site. 
      Connections (edges) in networks represent geographic distances (measured       to nearest01 m) between all lizards captured at each site (nodes, n ≥       28per site). Geographic distances were calculated from the latitude and        longitude data using the program GEOGRAPHIC DISTANCE MATRIX       
      GENERATORversion 1.2.3 (Ersts 2013).
    'mite_narea' = total number of distinct body regions with mites
    'mite_load' = mite rank abundance scaled 0–6(rank in parenthesis): no    
    mites (0), 1–5 mites (1), 6–10 mites (2),11–15 mites (3), 16–20 mites (4),     21–25 mites (5) and >25 mites(6)

 'contests'
Contains data for multivariate analysis of variance of male social contest behaviors. 

 'strength_data'
contains data for analysis of strength and mean strength of edges in each sites' spatial network. 

 'NBdiet' - 'HBdiet'
Each contain site-specific diet (trophic position) and microhabitat use (perch type) data for analyses comparing tree lizard colour morph differences in ecology. 
  'trt' = treatment (burn frequency)
  'gps' = lizard territory location/id
  'svl' = body size
  'tcol' = color morph (blue, orange, yellow)
  'perch' = microhabitat use either trees or snags
  'tpos' = trophic position

  NBdiet = non-burned site
  LBdiet = low-frequency burn site
  HBdiet = high-frequency burn site


# import data
```{r wood ants}
path_in <- "./data/LattanzioMiles_tree lizards/"
path_out <- "./analysis"  # where do we want to save the networks?


library(readxl)
morph_networks <- read_excel("../data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
    sheet = "morph_networks")
View(morph_networks)

contests <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                sheet = "contests")

strength <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                      sheet = "strength_data")

NBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "NBdiet")

LBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "LBdiet")

HBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "HBdiet")

```



# number of networks: spatial networks 
- condition/treatment: burn frequency sites (3)
- color morphs: blue, orange, yellow (3)

nodes: geographic locations of each lizards (lat, long)
lizard ids/territory names:
  'gps' in morph_networks and diets files 
  'id' in contest file

Connections (edges) in networks represent geographic distances (measured to nearest 0.1 m) between all lizards captured at each site (nodes, n ≥ 28per site)
  >50 m   edge weight = 0
  40-50 m             = 1
  30-40m              = 2
  20-30m              = 3
  10-20m              = 4
  0-10m               = 5


undirected networks


# get data ready

```{r}


```



get distance between pairs of tree lizards
```{r}

networkdata<- morph_networks %>% select(trt, gps, tcol, lat, long)

# distance per colony --> could be done with a loop too
unique(networkdata$trt)

#### NB ####
NB<-filter(networkdata, trt=="NB") %>% select(gps, long, lat)

CP <- function(NB) {
            Did = as.data.frame(t(combn(NB$gps, 2)))
            Z <- merge(Did, NB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, NB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_NB<- CP(NB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   

edge_distance_NB<- coord_NB%>%
  rowwise() %>%
  mutate(
         distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
         trt="NB")

edge_distance_NB$weight[edge_distance_NB$distance>50]<-0


## convert to network format
  graph_NB<-graph_from_data_frame(edge_distance_NB, directed = FALSE)
graph_NB
  
edge.attributes(graph_NB)$weight <- edge_distance_NB$weight

 #plot 
plot(graph_NB)



#### LB ####
LB<-filter(networkdata, trt=="LB") %>% select(gps, long, lat)

CP <- function(LB) {
            Did = as.data.frame(t(combn(LB$gps, 2)))
            Z <- merge(Did, LB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, LB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_LB<- CP(LB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   
edge_distance_LB<- coord_LB%>%
  rowwise() %>%
  mutate(
         distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
         trt="LB")

edge_distance_LB$weight[edge_distance_LB$distance>50]<-0


## convert to network format
  graph_LB<-graph_from_data_frame(edge_distance_LB, directed = FALSE)
graph 
  
 #plot 
plot(graph_LB)




#### HB ####
HB<-filter(networkdata, trt=="HB") %>% select(gps, long, lat)

CP <- function(HB) {
            Did = as.data.frame(t(combn(HB$gps, 2)))
            Z <- merge(Did, HB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, HB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_HB<- CP(HB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   
edge_distance_HB<- coord_HB%>%
  rowwise() %>%
  mutate(
    distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
    trt="HB")

edge_distance_HB$weight[edge_distance_HB$distance>50]<-0


## convert to network format
  graph_HB<-graph_from_data_frame(edge_distance_LB, directed = FALSE)
graph 
  
 #plot 
plot(graph_HB)


#### combine datasets
edge_distance <- bind_rows(edge_distance_NB, edge_distance_LB, edge_distance_HB) %>% select(trt, id1, id2, distance, weight)
glimpse(edge_distance)

```

# loop to get network and network metrics
```{r}

head(edge_distance)
trt<-unique(networkdata$trt)


 all_nets = list()  # empty list for networks
 
 #make empty dataframe to fill
summary <- data.frame(focal.m=character(), 
                      n_nodes=numeric(), 
                      density=numeric(), 
                      eigenvector_centrality=numeric(),
                      betweenness_centrality=numeric(),
                      normal_ei=numeric())
                     #apl=numeric(),
                     #ei=numeric())
                     #)
                                    
  i=1

for (i in 1:length(trt)) {
  # as a check and to know where in the loop we are 
  focal <- trt[i]
  print(focal) 
  
  # subset the dataframe per combi
  focal.data <- subset(edge_distance, trt==focal)
  head(focal.data)
  
  ## convert to network format
  graph<-graph_from_data_frame(focal.data, directed = FALSE)
  graph 
  
  edge.attributes(graph)$weight <- focal.data$weight
  
 #plot 
  plot(graph, 
       edge.width=E(graph)$weight*1,
       #vertex.color="grey",
       vertex.label="",
       edge.color=alpha("black", 1),
       vertex.size=20,
       layout=layout_in_circle,
       main=focal)

  graph <- as.list(graph)
  names(graph) <- focal
  all_nets[[focal]] <- graph
  
  
  # number of nodes
  n_nodes <- length(V(graph))
  
  ## individual-level 

# betweenness centrality
betweenness_centrality <- mean(igraph::betweenness(graph, directed = F,
                                      weights = E(graph)$weight))

# eigenvector centrality
eigen <- eigen_centrality(graph, directed = F, weights = E(graph)$weight)
ec <- eigen$vector
ec.value <- eigen$value
eigenvector_centrality<-mean(igraph::eigen_centrality(graph, directed = F)$vector)


## group-level 

# density =  number of observed dyads/total possible dyads
density<-igraph::edge_density(graph)
#density<-ecount(graph)/(vcount(graph)*(vcount(graph)-1)) #for a directed network

# average path length
apl <- igraph::mean_distance(graph) #average.path.length(graph.ref.crowd) 


# efficiency
library(einet)
ei <- effective_information(graph, effectiveness = FALSE)
normal_ei <- ei/log2(n_nodes) # normalized value to control for network size


#Find proportion unknown relationships, a measure of sparseness
#prunk <- EloRating::prunk(matrix)
#prunk.pu <- as.numeric(prunk[1])
#prunk.dyads <- as.numeric(prunk[2])

## pool all data
pool_metrics<- cbind.data.frame(focal.m,
                      #n, 
                      n_nodes,
                      density, 
                      eigenvector_centrality, 
                      betweenness_centrality,
                      #ec.value,
                      #apl,
                      normal_ei)

summary <- rbind(summary, pool_metrics)
                        
}
 
all_treelizards_networks<-all_nets


```


```{r}
# compile output and save 
head(summary)

network_output <- summary %>% 
  #tibble::rownames_to_column() %>% rename(nodeID=rowname)
  #separate(focal.m, c("replicate", "treatment")) %>%
  rename(treatment=trt) %>%
  mutate(study="LattanzioMiles", species="tree lizards", replicate="") %>%
  select(study, species, treatment, replicate, everything())
head(network_output)

write.csv(network_output, file = "network_output_treelizards.csv")
 
```


