---
title: "import_networks"
author: "Annemarie van der Marel"
date: "11/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(igraph)
library(ggplot2)
library(einet)
library(EloRating)
library(geosphere)
```

```{r}
#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}
```



# Import tree lizards networks
Lattanzio & Miles 2014

It's not a before and after a perturbation. Instead, spatial networks are compared between different sites differing in the frequency of a disturbance. 
We could use the non-burned site as control (before treatment) and high burn site as after treatment.  

 'morph_networks'
Contains data for morphology, body condition (svl), parasite load, bite mark, perch use frequency, network construction (lat, long data) and analysis (degree, eigenvector centrality data), and morph frequency analyses. 

    'lat' and 'long' used to estimate spatial network at each site. 
      Connections (edges) in networks represent geographic distances (measured       to nearest01 m) between all lizards captured at each site (nodes, n ≥       28per site). Geographic distances were calculated from the latitude and        longitude data using the program GEOGRAPHIC DISTANCE MATRIX       
      GENERATORversion 1.2.3 (Ersts 2013).
    'mite_narea' = total number of distinct body regions with mites
    'mite_load' = mite rank abundance scaled 0–6(rank in parenthesis): no    
    mites (0), 1–5 mites (1), 6–10 mites (2),11–15 mites (3), 16–20 mites (4),     21–25 mites (5) and >25 mites(6)

 'contests'
Contains data for multivariate analysis of variance of male social contest behaviors. 

 'strength_data'
contains data for analysis of strength and mean strength of edges in each sites' spatial network. 

 'NBdiet' - 'HBdiet'
Each contain site-specific diet (trophic position) and microhabitat use (perch type) data for analyses comparing tree lizard colour morph differences in ecology. 
  'trt' = treatment (burn frequency)
  'gps' = lizard territory location/id
  'svl' = body size
  'tcol' = color morph (blue, orange, yellow)
  'perch' = microhabitat use either trees or snags
  'tpos' = trophic position

  NBdiet = non-burned site
  LBdiet = low-frequency burn site
  HBdiet = high-frequency burn site


# import data
```{r wood ants}
path_in <- "./data/LattanzioMiles_tree lizards/"
path_out <- "./analysis"  # where do we want to save the networks?


library(readxl)
morph_networks <- read_excel("../data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
    sheet = "morph_networks")
View(morph_networks)

contests <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                sheet = "contests")

strength <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                      sheet = "strength_data")

NBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "NBdiet")

LBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "LBdiet")

HBdiet <- read_excel("data/LattanzioMiles_tree lizards/LattanzioMilesJAEdata.xlsx", 
                                     sheet = "HBdiet")

```



# number of networks: spatial networks 
- condition/treatment: burn frequency sites (3)
- color morphs: blue, orange, yellow (3)

nodes: geographic locations of each lizards (lat, long)
lizard ids/territory names:
  'gps' in morph_networks and diets files 
  'id' in contest file

Connections (edges) in networks represent geographic distances (measured to nearest 0.1 m) between all lizards captured at each site (nodes, n ≥ 28per site)
  >50 m   edge weight = 0
  40-50 m             = 1
  30-40m              = 2
  20-30m              = 3
  10-20m              = 4
  0-10m               = 5


undirected networks


# get data ready

```{r}


```



get distance between pairs of tree lizards
```{r}

networkdata<- morph_networks %>% select(trt, gps, tcol, lat, long)

# distance per colony --> could be done with a loop too
unique(networkdata$trt)

#### NB ####
NB<-filter(networkdata, trt=="NB") %>% select(gps, long, lat)

CP <- function(NB) {
            Did = as.data.frame(t(combn(NB$gps, 2)))
            Z <- merge(Did, NB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, NB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_NB<- CP(NB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   

edge_distance_NB<- coord_NB%>%
  rowwise() %>%
  mutate(
         distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
         trt="NB")

edge_distance_NB$weight[edge_distance_NB$distance>50]<-0

# filter out distance > 50 m
edge_distance_NB_filt <- edge_distance_NB %>% filter(weight!=0)

## convert to network format
  graph_NB<-graph_from_data_frame(edge_distance_NB_filt, directed = FALSE)
graph_NB
  
edge.attributes(graph_NB)$weight <- edge_distance_NB_filt$weight

 #plot 
plot(graph_NB)



#### LB ####
LB<-filter(networkdata, trt=="LB") %>% select(gps, long, lat)

CP <- function(LB) {
            Did = as.data.frame(t(combn(LB$gps, 2)))
            Z <- merge(Did, LB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, LB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_LB<- CP(LB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   
edge_distance_LB<- coord_LB%>%
  rowwise() %>%
  mutate(
         distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
         trt="LB")

edge_distance_LB$weight[edge_distance_LB$distance>50]<-0

# filter out distance > 50 m
edge_distance_LB_filt <- edge_distance_LB %>% filter(weight!=0)

## convert to network format
  graph_LB<-graph_from_data_frame(edge_distance_LB_filt, directed = FALSE)
graph_LB
  
edge.attributes(graph_LB)$weight <- edge_distance_LB_filt$weight




#### HB ####
HB<-filter(networkdata, trt=="HB") %>% select(gps, long, lat)

CP <- function(HB) {
            Did = as.data.frame(t(combn(HB$gps, 2)))
            Z <- merge(Did, HB, by.x = "V2", by.y = "gps")
            Z <- merge(Z, HB, by.x = "V1", by.y = "gps")
            return(Z)
      }                  
        
coord_HB<- CP(HB) %>% rename(id1=V1, id2=V2, long1=long.y, lat1=lat.y,
                                      long2=long.x, lat2=lat.x)
   
edge_distance_HB<- coord_HB%>%
  rowwise() %>%
  mutate(
    distance = distHaversine(cbind(long1,lat1),cbind(long2,lat2)),
         weight=if_else(distance>50, 0, 
                        if_else(distance>40, 1, 
                                if_else(distance>30, 2,
                                       if_else(distance>20, 3,
                                               if_else(distance>10, 4,
                                                       if_else(distance>0, 5, 5)))))),
    trt="HB")

edge_distance_HB$weight[edge_distance_HB$distance>50]<-0

# filter out distance > 50 m
edge_distance_HB_filt <- edge_distance_HB %>% filter(weight!=0)

## convert to network format
  graph_HB<-graph_from_data_frame(edge_distance_HB_filt, directed = FALSE)
graph_HB
  
edge.attributes(graph_HB)$weight <- edge_distance_HB_filt$weight

plot(graph_HB)

#### combine datasets
edge_distance <- bind_rows(edge_distance_NB, edge_distance_LB, edge_distance_HB) %>% select(trt, id1, id2, distance, weight)
glimpse(edge_distance)

# filter out edge weights of 0 
edge_distance_filt <- bind_rows(edge_distance_NB_filt, edge_distance_LB_filt, edge_distance_HB_filt) %>% select(trt, id1, id2, distance, weight)
glimpse(edge_distance_filt)


```

# loop to get network and network metrics
```{r}

head(edge_distance)
head(edge_distance_filt) # filtered dataset 0 edge weights
trt<-unique(networkdata$trt)

# set data
#data<-edge_distance %>% select(id1, id2, trt, distance, weight)
data<-edge_distance_filt %>% select(id1, id2, trt, distance, weight)


 all_nets = list()  # empty list for networks
 
 #make empty dataframe to fill
node_summary <- data.frame(focal=character(), 
                      degree=numeric(), 
                      eigenvector_centrality=numeric(),
                      betweenness_centrality=numeric())
 
 
network_summary <- data.frame(focal=character(), 
                      n_nodes=numeric(), 
                      density=numeric(), 
                      normal_ei=numeric())
                     #apl=numeric(),
                     #ei=numeric())
                     #)         
  i=1

for (i in 1:length(trt)) {
  # as a check and to know where in the loop we are 
  focal <- trt[i]
  print(focal) 
  
  # subset the dataframe per combi
  focal.data <- subset(data, trt==focal)
  head(focal.data)
  
  ## convert to network format
  graph<-graph_from_data_frame(focal.data, directed = F)
graph 
  
 #plot 
  plot(graph, 
       edge.width=E(graph)$weight,
       vertex.color="grey",
       vertex.label="",
       edge.color=alpha("black", 1),
       vertex.size=20,
       #layout=layout_in_circle,
       main=focal)


#### individual-level : Betweenness Centrality, Eigenvector Centrality, and Degree ###

  # degree
  degree <- igraph::degree(graph)
  
  # betweenness centrality -> standardized
  between <- igraph::betweenness(graph, directed = F,
                                      weights = E(graph)$weight)
  betweenness_centrality <- (between-min(between))/(max(between)-min(between))
  
  
  
# eigenvector centrality
eigenvector_centrality<-igraph::eigen_centrality(graph, directed = F)$vector


# pool individual
pool_id <- cbind.data.frame(focal,
                            degree,
                            betweenness_centrality, 
                            eigenvector_centrality)

node_summary <- rbind(node_summary, pool_id)
                        


#### network level (N Nodes, Density, Effective Information) ####

# number of nodes
  n_nodes <- length(V(graph))

# density =  number of observed dyads/total possible dyads
#density<-ecount(graph)/(vcount(graph)*(vcount(graph)-1)) #for a directed network
density<-igraph::edge_density(graph)

# average path length
apl <- igraph::mean_distance(graph) #average.path.length(graph.ref.crowd) 

# efficiency
library(einet)
ei <- effective_information(graph, effectiveness = FALSE)
normal_ei <- ei/log2(n_nodes) # normalized value to control for network size

# betweenness_centrality <- mean(igraph::betweenness(graph, directed = F, weights = E(graph)$weight))
  # eigenvector_centrality<-mean(igraph::eigen_centrality(graph, directed = F)$vector)

#Find proportion unknown relationships, a measure of sparseness
#prunk <- EloRating::prunk(matrix)
#prunk.pu <- as.numeric(prunk[1])
#prunk.dyads <- as.numeric(prunk[2])

## pool network data
pool_network<- cbind.data.frame(focal,
                      n_nodes,
                      density, 
                      normal_ei)

network_summary <- rbind(network_summary, pool_network)
     
# pool networks
  graph <- as.list(graph)
  names(graph) <- focal
  all_nets[[focal]] <- graph

                        
}
 
all_treelizards_networks<-all_nets


```

```{r}
# community detection

ceb <- cluster_edge_betweenness(graph) 
dendPlot(ceb, mode="hclust")
plot(ceb, graph) 

clp <- cluster_label_prop(graph)
plot(clp, graph)


```



```{r}
# compile output and save 
head(node_summary)

node_output<- node_summary %>% 
  tibble::rownames_to_column() %>% rename(nodeID=rowname) %>%
  #separate(focal.m, c("replicate", "treatment")) %>%
  rename(treatment=focal) %>%
  mutate(study="LattanzioMiles", species="tree lizards", replicate="") %>%
  select(study, species, treatment, replicate, everything())
node_output$treatment<- factor(node_output$treatment, levels=c("NB", "LB", "HB"), labels=c("no burn", "low burn", "high burn"))
head(node_output)

write.csv(node_output, file = "../analysis_results/node_output_treelizards_0weightexcl.csv")


# output network data
head(network_summary)

network_output <- network_summary %>% 
  #tibble::rownames_to_column() %>% rename(nodeID=rowname)
  #separate(focal.m, c("replicate", "treatment")) %>%
  rename(treatment=focal) %>%
  mutate(study="LattanzioMiles", species="tree lizards", replicate="") %>%
  select(study, species, treatment, replicate, everything()) 
  
  network_output$treatment<- factor(network_output$treatment, levels=c("NB", "LB", "HB"), labels=c("no burn", "low burn", "high burn"))
head(network_output)

write.csv(network_output, file = "../analysis_results/network_output_treelizards_0weightexcl.csv")
 

# save network edgelist or matrix as csv
head(edge_distance)
glimpse(edge_distance)
write.csv(edge_distance, file = "../analysis_results/edgelist_treelizards.csv")


head(edge_distance_filt)
glimpse(edge_distance_filt)
write.csv(edge_distance_filt, file = "../analysis_results/edgelist_treelizards_0weightexcl.csv")

```


